Refunctionalization of Abstracting Abstract Machines (Functional Pearl)

### 2018-03-17 paper.tex
intro

But in terms of their abstract semantic artifacts, 
the relations between small-step abstract abstract machines and big-step abstract 
definitional interpreters, and the question that how can we interdrive each other is not
well studied.
The fundamental difference between concrete semantics artifacts and abstract semantics 
artifacts is non-determinism. In addition to insure termination, abstract semantics 
artifacts usually equipped with a cache of explored state that is guaranteed to reach 
the least fixed-point eventually.
In this functional peral, we practically answer the question and relate AAM and ADI 
by presenting a series of syntactical transformations on the program from
small-step abstract abstract machines to big-step abstract definitional interpreters.

@@

\note{rest of intro goes here}

@@

\begin{itemize}
\item We show how \textbf{linearization} (Section~\ref{linear}) can be used to...
\item In Section~\ref{fusing}, we then present the \textbf{fusing} transformation, and present...
\item Section~\ref{disen} discusses \textbf{disentangling}, which shows...
\item \textbf{Refunctionalization} is shown in Section~\ref{refunc}, and brings a new understanding 
  of pushdown control-flow analysis. In this section, we present an abstract interpreter obtained 
  from refunctionalizing a small-step abstracting abstract machine which naturally exhibits the 
  pushdown control-flow property (as refunctionalization requires abstract execution to be sequentialized).
\item The last transformation we show is that of transforming a definitional abstract interpreter 
  to a \textbf{direct-style} interpreter (Section~\ref{directstyle}).

@@

These transformations are used throughout the paper, with refunctionalization and
defunctionalization of abstract interpreters playing important roles for the call stack of
the analyzed language. All of this insight allows the revelation of
precisely \emph{why} small-step AAM does not have the pushdown control-flow property
(without extra effort), as well as explains why ADI naturally extends pushdown control-flow
property from its defining language.

@@

1) The current control expression is @Let@,
and then the result of @App(f, ae)@ will be bound to variable @x@.
Firstly we retrieve the set of closures that @f@ may represent.
For each closure in the set, we do almost the same as what we did in the concrete
CESK machines. But note that since the continuation is allocated on store (@kstore@),
a new continuation address @new_kaddr@ is constructed, and a new frame
@Frame(x, e, env)@ paired with the current continuation address @kaddr@
is joined into @new_kaddr@. Finally, a list of successors states is generated.

2) An atomic expression @ae@ sits on the control position of the state.
The value of @ae@ is being returned to its caller.
We dereference the continuation address @kaddr@ and obtain a set of
continuations @conts@.
For each continuation in the set, we construct an environment based on the
environment of the frame (@env_f@), and bind @x@ to a newly created binding
address @baddr@; also update the store with @baddr@ and
the values that @ae@ represents. In every generated states,
the control becomes the expression @e@ in the frame,
and as we can tell from the name, continuation address @f_kaddr@
also comes from the frame.

>>>>

But in terms of their abstract semantic artifacts, 
the relations between small-step abstract abstract machines and big-step abstract 
definitional interpreters, and the question that how can we interdrive each other is not
well studied.
The fundamental difference between concrete semantics artifacts and abstract semantics 
artifacts is non-determinism. In addition to insure termination, abstract semantics 
artifacts are usually equipped with a cache of explored state that is guaranteed to reach 
the least fixed-point eventually.
In this functional peral, we practically answer the question and relate AAM and ADI 
by presenting a series of syntactical transformations on the program from
small-step abstract abstract machines to big-step abstract definitional interpreters.

In addition, the abstract abstract machines with unbounded stack naturally corresponds 
to abstract definitional interpreters. We show after refunctionalizing an AAM with unbounded
stack and with proper caching algorithm, the pushdown control-flow analysis can be obtained.

@@

@@

\begin{itemize}
\item We show \textbf{linearization} in Section~\ref{linear}. By expressing the 
  non-deterministic choices as first-order data type, we linearize the execution 
  of abstract abstract machines, therefore the transition of machine state becomes deterministic.
    Notably, this introduces anther layer of controls that also will be refunctionalized later.

\item In Section~\ref{fusing}, we then present the \textbf{fusing} transformation. 
  The fusing transformation simply combines the single-step function @step@ and the driving
  function into one, but keeps all the machine state representation.

\item Section~\ref{disen} discusses \textbf{disentangling}, which disassemble the fused AAM to be
  several individual functions, and each function handles one data type represented continuation.

\item \textbf{Refunctionalization} is shown in Section~\ref{refunc} which sequentializes
  the order of abstract execution by higher-order functions.
  For clarity, we first present the refunctionalized AAM without caching. Then we adopt
    a different caching algorithm to guarantee the termination of abstract interpretation.
  In this section, we also review pushdown control-flow analysis and examine how computable 
    and precise call/return match is obtained through these transformations so far.

\item The last transformation we show is that of transforming the refunctionalized AAM
  to a \textbf{direct-style} interpreter (Section~\ref{directstyle}) by using delimited
    controls.

@@

These transformations are used throughout the paper, with refunctionalization and
defunctionalization of abstract interpreters playing important roles for the call stack of
the analyzed language. By refunctionalization, the call stack of the analyzed language
is blended into the call stack of the defining language. This provides another perspective to 
explain why \citeauthor{darais2017abstracting}'s abstract definitional interpreters 
is able to extend the the pushdown control-flow property from its defining language.

@@

\begin{itemize}
  \item The current control expression is @Let@,
and then the result of @App(f, ae)@ will be bound to variable @x@.
Firstly we retrieve the set of closures that @f@ may represent.
For each closure in the set, we do almost the same as what we did in the concrete
CESK machines. But note that since the continuation is allocated on store (@kstore@),
a new continuation address @new_kaddr@ is constructed, and a new frame
@Frame(x, e, env)@ paired with the current continuation address @kaddr@
is joined into @new_kaddr@. Finally, a list of successors states is generated.

  \item An atomic expression @ae@ sits on the control position of the state.
The value of @ae@ is being returned to its caller.
We dereference the continuation address @kaddr@ and obtain a set of
continuations @conts@.
For each continuation in the set, we construct an environment based on the
environment of the frame (@env_f@), and bind @x@ to a newly created binding
address @baddr@; also update the store with @baddr@ and
the values that @ae@ represents. In every generated states,
the control becomes the expression @e@ in the frame,
and as we can tell from the name, continuation address @f_kaddr@
also comes from the frame.
\end{itemize}
